---
title: "Cube Conundrum Part One"
author: "Aman Das"
date: 2023-12-2
---

## Introduction

It's my brother's birthday today. Let's get this cute possibility check done.

**TODO:**

-   Extract input file
-   organize the input into a data frame
-   complete the puzzle check
-   sum the valid game ids
-   output the result

```{r}
library(tidyverse)
```

## Read the input

```{r}
games = read_delim("input.txt", delim=":", col_names = c("game", "result"))
```

## Manipulate the games data

```{r}
games = games %>%                         # get the data
  mutate(                                 # manipulate it 
    game = game %>%                       # from the game id column
      str_extract("\\d+") %>%             # get the 1st numeral
      as.numeric(),                       # and convert to numeric
    red = result %>%                      # RED get results column
      str_extract_all( "\\d+ red") %>%    # get all the "000 red" sub strings as a vector
      str_extract_all("\\d+") %>%         # get numerals from each vector
      lapply( as.numeric) %>%             # convert each element of vector to numeric
      lapply( max) %>%                    # find max in each vector
      flatten_dbl(),                      # from list of vectors of max, to vector of max
    green = result %>%                    # GREEN
      str_extract_all( "\\d+ green") %>%
      str_extract_all("\\d+") %>%
      lapply( as.numeric) %>%
      lapply( max) %>%
      flatten_dbl(),
    blue = result %>%                     # BLUE
      str_extract_all( "\\d+ blue") %>%
      str_extract_all("\\d+") %>%
      lapply( as.numeric) %>%
      lapply( max) %>%
      flatten_dbl(),
    result = NULL
  )
```

## Check the conditions and sum the game numbers

```{r}
result = games %>%                        # get the tibble
  filter(                                 # filter rows
    red <= 12,                            # such that no. of max red cubes <= 12
    green <= 13,                          # no. of max green cubes <= 13
    blue <= 14                            # no. of max blue cubes <= 14
  ) %>%                                   # get the filtered tibble
  select( "game" ) %>%                    # get the game id column
  sum()                                   # add them all up
```

## Output the result

```{r}
write_lines(result, "output.txt")         # output
```

## Conclusion

I love and hate regex. Need to learn pipe operations.